---
title: "How to GWAS - ConGen 2022"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output: beamer_presentation
header-includes:
  - \usepackage{wrapfig}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is GWAS?

Genome Wide Association Study

- y = a + bx + g + e
- y: phenotype
- b: fixed effect
- x: the SNP predictor coded as 0, 1, 2
- g: the polygenic random effect (e.g. Kinship matrix)
- e: the error term

## Soay Sheep Horns RXFP2 genotype

Johnson et al. 2011 Mol Ecol, Johnson et al. 2013 Nature
![](Johnson_Phenotypes.pdf){width=45%} ![](Johnson_Manhatten_plot.pdf){width=65%}

- 486 sheep
- 35 831 SNPs
- one huge effect locus, explains 76% of the variation in horn size


## How often do people find huge effect loci using GWAS?

 ![](Margres_2018_Tazzies.pdf){width=50%}
- Darwin's finches
- sticklebacks
- atlantic salmon
- Tasmanian devils face cancer 5 SNPs explain ~61% of variation in female survival

## What are the expectations for a GWAS?

![](Silva_flycatcher.pdf){width=90%}

- No significant SNPs for morphological traits
- 3 SNPs explaining ~3% of the variation in clutch size

Silva et al. 2017 Heredity, Husby et al. 2014 Proc B

## What are the expectations for a GWAS?

- Rockman 2012 - All that's gold does not glitter.


## Some considerations for GWAS

\begin{minipage}[t]{.45\linewidth}\centering


- heritability of the trait

- genetic architecture of the trait (unknown!)

- Linkage Disequilibrium

- number of individuals sampled


\end{minipage}
\begin{minipage}[t]{.45\linewidth}
```{r, echo=FALSE, fig.align='right', out.height='120%'}
knitr::include_graphics("Gompert2017_Figure5.pdf")
```
\end{minipage}
                          
Kardos 2015 MER, Gompert 2017 Mol Ecol

## Some solutions

- Sample more individuals!

- replicated samples of individuals (RepeatABEL, RIP)

- Admixture mapping instead of within population GWAS

- sample in closed populations (but know this limits generalizability)

## What is Admixture mapping?

## Deer Case study

## So what are we doing today?

- Simulated GWAS, using GEMMA

