---
title: "How to GWAS - ConGen 2022"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is GWAS?

Genome Wide Association Study

- y = a + bx + g + e
- y: phenotype
- b: fixed effect
- x: the SNP predictor coded as 0, 1, 2
- g: the polygenic random effect (e.g. Kinship matrix)
- e: the error term

## Soay Sheep Horns RXFP2 genotype

Johnson et al. 2011 Mol Ecol, Johnson et al. 2013 Nature
![](Johnson_Phenotypes.pdf){width=45%} ![](Johnson_Manhatten_plot.pdf){width=65%}

- 486 sheep
- 35 831 SNPs
- one huge effect locus, explains 76% of the variation in horn size


## What are the expectations for a GWAS?

![](Silva_flycatcher.pdf){width=90%}

- No significant SNPs for morphological traits
- 3 SNPs explaining ~3% of the variation in clutch size

Silva et al. 2017 Heredity, Husby et al. 2014 Proc B

## Some considerations for GWAS

- heritability of the trait!
- Linkage Disequilibrium
- number of individuals sampled


## Slide with Plot

```{r pressure}
plot(pressure)
```

